Inference of seasonal forcings on reproduction numbers during epidemics

### 2020-04-18 PaperDraft.Rmd
add on discussion

For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$. This method is outlined in detail in Wallinga and Teunis (2004)[8].

@@

[What methods/models are best]
[What condition(s) does each method/model work well/poorly]
[Conclusion about dataset]

@@

## Discussion

>>>>

For each province, we take a 7 day rolling sum on the number of new cases each day, and feed the data into the parametric SI model, which is given distribution $w(\tau)$ From there, the model uses a likelihood based approach to estiamte R at each discrete timestep, computing the likelihood that a person at timestep $i$ is infected by a person infected at timestep $j$, and summing the likelihood for each $j$ across all its infectees to compute the $R$ value at timestep $j$. Thus, for each province, we produce a daily time series estimate for $R$ on each day. This method is outlined in detail in Wallinga and Teunis (2004)[8].

For each province, day pair, we have $\bar{x} = (x_1, x_2, x_3, x_4), y$ where:

- $x_1$: standardized GDP per capita
- $x_2$: standardized population density
- $x_3$: standardized air temperature
- $x_4$: standardized relative humidity
- $y$: logarithm of estimated R

We then perform ordinary least squares regression to find $\bar{\beta}$:
$$
y = \beta_0 + \beta_1x_1 + \beta_2x_2 + \beta_3 x_3 + \beta_4 x_4 + \varepsilon
$$

@@

@@

### Interpretation

While the overall results are not surprising, some caution is important in the
interpretation of these results. There are at least three sources of uncertainty
not captured in the analysis:

- Uncertainty about the serial interval distribution (our method assumes the
  serial interval distribution is known)
- Uncertainty about timeliness of data (our model assumes cases are reported on
  the date of infection; in reality there is a significant delay, which the model
  does not attempt to adjust for)
- Uncertainty of the completeness of the data (our method assumes all cases are
  reported, which may not be true, as the high spikes in estimated $R$ are
  unlikely to be ground truth; a more likely explanation is that early cases went
  undetected for a long time)
- Uncertainty about the accuracy and currency of data (our model assumes that
  temperature, relative humidity, GDP per capita, and population data are local
  and current, but in reality the demographic data is from 2017 and the weather
  data is measured by the closest station that may be in some cases not even
  located in the province)

Hence, it is likely that the uncertainty in posterior estimations of $\beta$ by
the model is understated. While the model provides some evidence that
temperature is effective at reducing spread, we do not believe this evidence to
be as strong as claimed.

## Discussion
Our model suggests that there is some weak evidence that with increasing temperature, the reproductive number R decreases. The effect of relative humidity on the reproductive number is much less conclusive. As we only used data from Italy in our model building, there is a limited range of weather information than we would like to have.
